Manuscript for foodborne-outbreak Project

### 2023-02-24 products/manuscript/Manuscript.qmd
part 2 update

*Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement.*

Using the datasets previously identified above, we analyzed the structure of each dataset and how many NA values were present for each column in the datasets. Afterwards, we selected the variables containing less than 80% NA values for each dataset, which included our predictors and outcomes of interest. We then filtered the dataset designated for analysis to include only foodborne disease outbreaks. For both datasets, we separated rows to address coinfections individually, and either created or modified the IFSAC Category variable to categorize salads, desserts, and poultry. We also created a variable for pathogen type (virus, bacteria, parasite, or toxin) and species classification.

Post-preprocessing, we performed an exploratory data analysis to view general relationships between the predictors: time (year), pathogen type and etiology, state, food type, and health outcomes of interest: hospitalizations and illnesses.

We will be building regression models using these predictors and health outcomes as well; we

1.  Data wrangling: handling the missing values and cleaning inconsistent data.
2.  Selecting the interested predictors and outcomes to plot their relationships.
3.  If the graphs indicate any significant or interesting trend, we will conduct a general linear model and omnibus test.
4.  Post-hoc test to identify the significant difference between levels.
5.  Mapping cases with different predictors across the states.
6.  Applying Machine learning approach to test the model accuracy.

@@

*As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next.*

@@

## Data import and cleaning

@@

*Write code that reads in the file and cleans it so it's ready for analysis. Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts. If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g. as supplement in a paper). All materials, including files that contain code, should be commented well so everyone can follow along.*

```{r}
#The detailed cleaning process can be found in the file located in ("code","processing","processingfile_v1.qmd")

```

## Statistical analysis

@@

## Exploratory/Descriptive analysis

>>>>

@@

@@

## Data Importation and Cleaning

Data was imported into RStudio using the tidyverse and readxl packages. Using the datasets identified above, we analyzed the structure of each dataset and how many NA values were present for each column in the datasets. Afterwards, we selected the variables containing less than 80% NA values for each dataset, which included our predictors and outcomes of interest. We then filtered the dataset designated for analysis to include only foodborne disease outbreaks. For both datasets, we separated rows to address coinfections individually, and either created or modified the IFSAC Category variable to categorize salads, desserts, and poultry. We also created a variable for pathogen type (virus, bacteria, parasite, or toxin) and species classification. We also renamed setting/location (private residence, restaurant, etc.) of outbreaks for consistency.

@@

## Statistical Analysis
Post-preprocessing, we performed an exploratory data analysis to view general relationships between the predictors: time (year), pathogen type and etiology, state, food type, and health outcomes of interest: hospitalizations and illnesses.

We will be building regression models using these predictors and health outcomes as well; we intend to implement machine learning to train a model to predict health outcomes using the Kaggle dataset for the CDC NORS data.

@@

## Exploratory/Descriptive Analysis
